---
title: "Forest Type to Assessment Area Crosswalk"
author: "Alec Trusty"
date: "1/27/2022"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, include=FALSE}
library(sf)
library(rjson)
library(dplyr)
library(data.table)
library(knitr)

# Set paths
raster_dir <- '/home/atrusty/Projects/Heatmap/SilviaTerra/dendroBasemap/'
data_dir <- '/home/atrusty/github/ftype-to-assessment/data/'
```

# Forest Type to Assessment Area Crosswalk

This document describes the steps for generating a forest type to assessment area crosswalk using CAR supersections, FIA forest types, ARB species-to-forest, and Carbon plan assessment area and common practice. The resulting dataset includes a many-to-many mapping of forest types to assessment areas and common practice values. This crosswalk is later applied to forest type and supersection raster data to generate raster datasets of assessment area and common practice values.

### Data loading

First we read in the supersections vector data for the CONUS supplied by the Carbon Action Reserve (CAR):

```{r Read SuperSections, echo=FALSE, include=FALSE}
# Read in supersection shapefile
supersect_sf <- st_read(paste0(data_dir, 'CAR_supersections/CAR_supersections_coded.shp'))
supersect_sf$ss_code <- as.character(supersect_sf$ss_code)
```

```{r}
kable(head(supersect_sf))
```

Next we read in a mapping of FIA Forest Types to common names:

```{r Read ForestType, echo=FALSE, include=FALSE}
# Read in forest type to common name and format as data table
fortypcmn_json <- fromJSON(file = paste0(data_dir, 'fia_fortyp_to_common_name.json'))
fortypcmn_dt <- fortypcmn_json %>% data.table()
names(fortypcmn_dt) <- as.character("ftype_name")
fortypcmn_dt$ftype_code <- lapply(seq_along(fortypcmn_json), function(i) names(fortypcmn_json)[[i]]) %>% unlist()
```

```{r}
kable(head(fortypcmn_dt))
```

As well as species-to-forest type codes from ARB:
```{r Read Species-to-Forest, echo=FALSE, include=FALSE}
# Read in species-forest type codes and format as a data table
fortypcds_json <- fromJSON(file = paste0(data_dir, "arb_fortypcds.json")) # 348
fortypcds_dt <- fortypcds_json %>% data.table()
names(fortypcds_dt) <- as.character("ftype_list")
fortypcds_dt$aa_code <- lapply(seq_along(fortypcds_json), function(i) names(fortypcds_json)[[i]]) %>% unlist()
```

```{r}
kable(head(fortypcds_dt))
```

Lastly we read in the assessment area & common practice data from Carbon Plan Paper (TODO: Add Link):

```{r, echo=FALSE, include=FALSE}
# Read in assessment area data 
aa_dt <- fread(paste0(data_dir,'2015_assessment_area_data_file.csv')) %>%
  .[, c("supersection", "assessment_area", "aa_code", "ss_code", "species", "site_class", "common_practice")]
```

```{r}
kable(head(aa_dt))
```

### Data Processing and Formatting

This next section shows the steps to wrangle all of these inputs into a crosswalk table that links all of the forest types and supersections to their corresponding assessment areas and common practice values.

Update assessment area names and format into a dataframe:

```{r}
# Update assessment area name
aa_dt$assessment_area <- lapply(aa_dt$assessment_area, function(x){
  ss_name <- aa_dt[assessment_area == x,]$supersection[[1]]
  aa_name <- aa_dt[assessment_area == x,]$assessment_area[[1]]
  ss_words <- strsplit(ss_name, " ") %>% unlist()
  aa_words <- strsplit(aa_name, " ") %>% unlist()

  aa_only <- aa_words[!(aa_words %in% ss_words)]
  paste0(aa_only, collapse = " ")
}) %>% unlist()

aa_uni_dt <- aa_dt[, list(
  ss_name= min(supersection),
  aa_name = min(assessment_area),
  species = unique(species),
  cp_lo = min(common_practice),
  cp_hi = max(common_practice), 
  ss_code = as.character(unique(ss_code)),
  aa_code = as.character(unique(aa_code))
), by=c("assessment_area", "species", "supersection")]

# WHAT ABOUT SITES WITH 'site_class'=='all'???
sc_all <- c("155", "156", "157", "158", "159", "160", "161")
aa_uni_dt <- aa_uni_dt[!(aa_uni_dt$aa_code %in% sc_all),]
```

```{r}
kable(head(aa_uni_dt))
```

Join forest type lists to assessment area table:

```{r}
# Join the forest type list to assessment data table
aa_ftype_dt <- aa_uni_dt[fortypcds_dt, on = 'aa_code', c("ss_name", "ss_code", "ftype_list", "aa_name", "species", "aa_code", "cp_lo", "cp_hi")]

# Break out the ftypes into individual rows
aa_ftype_expanded_dt <- lapply(aa_ftype_dt$aa_code, function(x)  {
  
  dt <- aa_ftype_dt[aa_ftype_dt$aa_code == x, ]
  
  dt_ftype_list <- dt$ftype_list[[1]]

  dt_expanded <- lapply(dt_ftype_list, function(y){
    dt_exp_y <- dt
    dt_exp_y$ftype_code <- as.character(y)
    dt_exp_y
  }) %>% rbindlist()

  dt_expanded
}) %>% rbindlist()
```

Assessment areas with forest type list, species, and common practice:

```{r}
kable(head(aa_ftype_dt))
```

```{r}
kable(head(aa_ftype_expanded_dt))
```

Update some forest type codes from FIA DB documentation (TODO: Include link to documentation)
```{r}
# Join the forest type common name to the data table
aa_ftype_expanded_dt <- fortypcmn_dt[aa_ftype_expanded_dt, on='ftype_code']
aa_ftype_expanded_dt$ftype_name <- as.character(aa_ftype_expanded_dt$ftype_name)

# Forest type 932 has been replaced with 933 in the most recent FIADB documentation,
# and 933 is fairly common in the CONUS ftype raster
aa_ftype_expanded_dt[ftype_code==932]$ftype_code <- 933

# Trim columns
cols2keep <- c("ftype_code", "ftype_name", "aa_code", "aa_name", "ss_code", "ss_name", "cp_lo", "cp_hi", "species")
aa_crosswalk_dt <- aa_ftype_expanded_dt[,cols2keep, with=FALSE]

# AT THIS POINT, the crosswalk is missing classes 203, 934 and 935 -- hardcode them
# Notes: 
# - There are only 3 pixels with class 203 (Big Cone Douglas Fir) which only occur in San Rafael and Los Padres forests in SoCal
# - 934 and 935 are both oak variety that occur in Sierra foothills
ftype_missing_dt <- fortypcmn_dt[ftype_code %in% c(203, 934, 935)]
ftype_missing_dt$aa_code <- c("294", "281", "281")# Calling Bigcone mixed conifer, and the live oaks Mixed Oak Woodland ( same as 'Canyon live oak / interior live oak)
aa_missing_dt <- aa_crosswalk_dt[aa_code %in% c(294, 281), !(names(aa_crosswalk_dt) %in% c("ftype_code", "ftype_name")), with=FALSE] %>% distinct()
ftype_addons_dt <- ftype_missing_dt[aa_missing_dt, on="aa_code"]

```

Join the updated reformatted fores types:
```{r, echo=FALSE, include=FALSE}
# Join the new data to the crosswalk
aa_crosswalk_dt <- rbind(aa_crosswalk_dt, ftype_addons_dt, fill=T)

# TODO: Maybe better to fix this at an earlier step
aa_crosswalk_dt$aa_name <- paste(aa_crosswalk_dt$ss_name, aa_crosswalk_dt$aa_name)
```

Final crosswalk table

```{r}
kable(head(aa_crosswalk_dt))
```

